\chapter{Proposed research plan}
As discussed in \cref{sec:challenges}, the study of additive manufacturing is a rather challenging task, mainly because industrial interests aimed to improve the AM technology are strongly related to issues that span over a wide length scale, from meso--scale to the atomistic level. Many phenomena that play a central role in the entire manufacturing process require the synergic collaboration of different approaches to obtain meaningful results.

We can summarize the previous statement in the following way: applications at the industrial level frequently encounters a series of metallurgical problems that firstly are to be fully understood to improve quality features of AM produced parts; on the other hand, AM processes are way too complicated to be directly studied by means of atomistic simulations and for this reason several phase field models have been developed to give more useful guidelines to industrial applications. However, in many cases even phase field models lack important knowledge of fundamental phenomena occurring, for example, during phase transitions or other basic physical properties. It is in the latter context that atomistic simulations can give a precious contribution. \Cref{tab:modeling_roadmap} is a sort of prospectus on the link between problems typically encountered at the industrial level, meso--scale study of AM and atomistic simulations. In this section on future research work we will explain a bit in detail how we want to study each of these topics.

\begin{table}[tb]
\centering
\caption{Summary table on frequent issues met on the industrial side and related computational approaches.}
\label{tab:modeling_roadmap}
\begin{tabularx}{\textwidth}{*{3}{X}}
\toprule
    \textbf{Metallurgical problems} & \textbf{Meso--scale approach} & \textbf{Atomistic parameters} \\
    \midrule
    Solute trapping & Redistribution coefficient & Interface mobility and diffusion matrix \\
    Gibbs--Marangoni effect & Convection, fluid flow simulation & Viscosity as a function of temperature\\
    Anisotropic growth (i.e.\ dendritic regime) & $\gamma$ dependence on orientation & Anisotropy of $\gamma$ \\
    Nucleation rate & Nucleation rate using classical nucleation theory and Avrami model & $\gamma$, enthalpy of melting, dependence of $\gamma$ on temperature and concentrations\\
    \bottomrule 
\end{tabularx}
\end{table}

During the first year of the PhD program, a considerable amount of time was devoted to getting familiar with the state--of--the--art about AM: its features as an innovative manufacturing process, the mechanisms presently used for processing metallic and related materials and a special focus on current challenges. The project is partially funded by several companies, that are interested to the application of AM to the energy sector (turbine blades)\footnote{The US company General Electric.} and to micro technology and jewelry (watch manufacturing)\footnote{Swatch and Rolex.}.

active in the field of industrial manufacturing, so the issues reported in \cref{tab:modeling_roadmap} are those of particular interest at the industrial level. Furthermore, this first part also included the study of most commonly used metal alloys for industrial applications of interest. 

A second task addressed during the first year involved the development and implementation of the revisited capillary fluctuation method with the aim of calculating the stiffness for different surfaces for an FCC system at the melting temperature. The comparison with results already present in literature showed a good agreement (see \cref{sec:results}).



\section{Topics of future work}

\subsection{Fundamental properties of real alloy systems}
The first future goal of the project will be to employ semi--empirical potentials and to model real alloy systems, namely Ni--Al--Cr and later Au--Ag--Cu. By means of molecular dynamics simulations, the aim is to extract useful informations on fundamental properties of the bulk liquid, like viscosity as a function of temperature and the so--called diffusion matrix. The latter plays the same role of the diffusion coefficient of a simple system, but in this case it takes into account the presence of more than one chemical species. All these properties do not require new method development and standard techniques of molecular simulation can be employed to extract them.

In this phase, much more emphasis will be given to the interatomic potential employed to model as accurately as possible these real systems. For metals and alloys, a well established choice involves the \textit{embedded atom method} (EAM) potentials, developed in 1983 by~\textcite{Daw1983EAM,Daw1984EAM} and subsequently improved by~\textcite{Finnis1984EAM} in 1984 and by~\textcite{Lee2000}, the latter to include second nearest neighbour interactions. EAM is the first choice for doing semi--empirical calculations in close--packed metals and related compounds such as alloys, because it combines the computational simplicity needed for large systems with a physical description of the underlying interactions that includes many--body effects, which are completely ignored in a pair--bond model.

The idea behind EAM is to write the total energy of a metal as a sum of energies obtained by inserting an atom into the local electron density produced by the remaining atoms of the system. Since we are dealing with charged particles, there is an additional pair term that takes into account electrostatic interaction. The expression of the atomic energy in the scheme of Finnis and Sinclair is of the form
\begin{equation}
    \label{eqn:EAM1}
    E_i= F_\alpha\left(\sum_{j\neq i} \rho_{\alpha\beta} (r_{ij}) \right) + \frac{1}{2} \sum_{j\neq i} \phi_{\alpha\beta}(r_{ij})
\end{equation}
Here $F_\alpha$ is the embedding energy, defined as the interaction of the atom $i$ of type $\alpha$ with the background electron gas; $\rho_{\alpha\beta}$ is the averaged atom electron density (a functional specific to the atomic types of both atoms) and $\phi$ is the pairwise term taking into account electrostatic interaction. The background density for each atom is determined by evaluating at its nucleus the superposition of atomic density tails from the other atoms.

When dealing with alloys, the EAM presents a practical advantage: the pairwise term involves only two atoms at a time and therefore can be calculated from the binary interactions involved in the ternary system in question. The embedding functions are defined separately for each of the components of the system and does not require any additional fitting beyond that done for pure components. Also the local electronic density depends only on the strength of each individual contributions of each of the three components. It follows that if EAM potentials are available for the three binary systems involved in a particular ternary, it is in theory possible to describe also the ternary system. The only requirement is that the interatomic interaction used for each of the pure components in the description of the two binaries involving that constituent be the same. However, even though in the original work of~\textcite{Daw1984EAM} were given mixing rules to obtain $\rho_{\alpha\beta}$ and $\phi_{\alpha\beta}$ for any alloy once known the corresponding functions for the elemental systems, it was soon made clear that this approximation was not enough to produce reliable and transferable potentials with the desired accuracy and the EAM was then revisited to include directly the fit to properties of the complex system of interest.

As an example of limitations of EAM potentials, \textcite{Grochola2005} in 2005 developed a new EAM potential for elemental Au. In \cref{tab:GoldEAM} is shown a comparison of some properties of Au predicted by several potentials as well as the available experimental data. It is evident that the predicting power of EAM is not able to provide an overall good accuracy on all the properties, being very precise for the cohesive energy but showing a considerable discrepancy on other fundamental quantities like the melting temperature. These deficiencies of the EAM stem from the method employed for fitting experimental data to the functional form of the potential and, from a theoretical point of view, from the fact that EAM contains approximations of the complicated many--body interactions which are the origin on many properties of condensed matter. For this reason, as we discuss in the next section, it 
\begin{table}[tb]
    \centering
    \caption{Comparison table between predicted values of EAM potentials for gold discussed (and referenced) in this section. The table is partially reproduced from Grochola et al~\cite{Grochola2005}. Units: elastic constants in \si{\giga\pascal}, energies in \si{eV} and lattice parameters in \si{\angstrom}.}
    \begin{tabular}{lSSSSS}
        \toprule
         \multirow{2}*{Property} & \multicolumn{4}{c}{Potential developep by} & \\
         \cmidrule(lr){2-5}
                & \text{Grochola} & \text{Ercolsssi~\cite{Ercolessi1988}} &  \text{Johnson~\cite{Jonhson1988:EAM}} & \text{Foiles~\cite{Foiles1986:EAMfcc}} & \text{Exp.} \\
         \midrule
          \text{Cohesive energy} & -3.924 & -3.78 & -3.930 & -3.927 & -3.93 \\
          \text{Lattice constant} & 4.0701 & 4.0704 & 4.0806 & 4.0805 & 4.07 \\
          \text{Bulk modulus} & 1.8026 & 1.8037 & 1.6987 & 1.6673 & 1.803 \\
          $c_{11}-c_{12}$ & 0.3207 & 0.5998 & 0.2687 & 0.2454 & 0.319 \\
          $c_{44}$ & 0.4594 & 0.5998 & 0.4069 & 0.4524 & 0.454 \\
          \text{Melting point (\si{\kelvin})} & 1159 & 1338 & 1053 & 1121 & 1337 \\
        \bottomrule
    \end{tabular}
    \label{tab:GoldEAM}
\end{table}

The first goal of the project will focus on modeling properties of undercooled melt of Ni--based superalloys, with compositions of specific alloys already employed in many applications\footnote{To the project are also participating companies (Heraeus, Oerlikon) that will directly supply these materials to experimental groups involved.}. Properties like viscosity and diffusion coefficients will be provided to a collaborating group at EMPA\footnote{Laboratory of Advanced Materials Processing, Dr.~Christian Leinenbach.} where phase field model are developed and used to study AM from a meso--scale point of view.



\subsection{Interface related properties}
During first--order phase transition such as nucleation and growth, interfacial properties play a central role. In particular, the interfacial free energy between the solid and the liquid phase (usually indicated with $\gamma_{sl}$) controls the barrier for nucleation of solid in an undercooled liquid and the crystal growth, which can present different regimes: planar, cellular and dendritic, the latter of particular interest for metals.

Despite its importance from both theoretical models and practical applications, an accurate calculation of $\gamma_{sl}$ is a rather complicated task even for the case of simple elements. Furthermore, on the experimental side, few techniques aimed at measuring this quantity are complicated by the very strict control on all experimental parameters that must be achieved to obtain accurate data.% One method, for example, involves recovering $\gamma_{sl}$ from nucleation--rate measurements: in this particular case, a major difficulty arises from the possible occurrence of heterogeneous nucleation from very low--concentration of impurities.

Several computational methods have been developed to calculate $\gamma_{sl}$, where a complete control on the experimental variables is possible. These methods includes the already discussed \emph{capillary fluctuation method}, different sorts of so--called \emph{cleaving methods} (CMs) and approaches based on \emph{classical nucleation theory} (CNT). All these schemes largely rely on molecular dynamics in conjunction with Monte Carlo simulations.

Recently, approaches of this kind coupled with the accelerated sampling technique of metadynamics have been presented: the key idea is to obtain $\gamma_{sl}$ from a free--energy map of the phase transition reconstructed by metadynamics~\cite{Angioletti-Uberti2010}, allowing also to investigate solidification and melting in out--of--equilibrium conditions~\cite{Cheng2015}, much more relevant for experiments and mesoscale models of solidification (including additive manufacturing processes).

Interface free energy is one among the interesting properties regarding the scenario of phase coexistence. It is worth mentioning that the anisotropy of $\gamma$ plays a central role when dealing with metals, because it is responsible of the dendritic growth regime. In addition to these, also the parameters related to the important phenomenon of nucleation are of great interest, because studying how the material behaves after the interaction with the laser source is a key factor to understand the microstructural evolution of the processed part and in turn its mechanical properties.


All these properties represent a rather challenging task from the molecular simulation point of view, mainly because they require specific methods to be employed in order to extract meaningful informations and because some properties may impose several constraints to the method used. One example concerns again $\gamma_{sl}$ and the CFM: as previously explained, this method has the drawback of requiring very large simulation cells with thousands of atoms to provide reliable results. Incidentally, the CFM does not measure directly the interface free energy but rather the stiffness, since the latter is one order of magnitude more anisotropic than $\gamma$ itself and can be determined with smaller errors.

An intrinsic technical problem concerning simulations of this kind is related to the potential used to model the interactions for these systems. Real alloys often consist of very complicated multi--component systems, with more than ten constituents: tackling the study of such systems with an approximate method like EAM would probably give useless results, not to mention that applying a method such as the CFM on such complex systems would mean performing simulations enormously complicated and with a very wide margin of error on any result.

An alternative and quite promising approach would be devoting a portion of time in developing more accurate and transferable potentials by means of powerful machine learning techniques such as Neural Networks. This kind of potentials are ``trained'' on a great number of relevant configurations of the system to model and they have shown to provide results with an accuracy close to that of \textit{ab initio} methods at a far lower computational cost (comparable to semi--empirical methods). One foreseeable limitation is related to the exponential complexity these potentials present when applied to study systems with too many degrees of freedom, such as ternary systems; unfortunately, this is the case of real alloys on Ni and Au on which we would be more interested. In conclusion, such an advanced approach as Neural Networks will have to be limited to the study of simpler systems that can serve as ``prototypes'' for complex alloys; at the same time, it can still provide fairly accurate informations on properties related to the coexistence of different phases (not only liquid and solid) in these materials.




\subsection{Timelines}